Unshuffling Permutations

### 2015-09-27 submission.tex
NP-hard proof
Fig 5

  It is immediate that H can be constructed in polynomial-time in $n$ and $k$.
  % We claim that $\sigma$ occurs in $\pi$ if and only if
  % there exists a containment-free perfect matching
  % $\mathcal{M}$ in $K_\mu$ such that
  % for any two disctinct edges
  % $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  % we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
  It can be shown (see Appendix) that
  that $\sigma$ occurs in $\pi$ if and only if
  there exists a containment-free perfect matching
  $\mathcal{M}$ in $K_\mu$ such that
  for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
  \qed
\end{proof}

@@

          \node [label] (y) at (17,14.5) {$y$};
          \node [label] (yp) at (10,-1.5) {$y'$};

@@

        \node [label] (y) at (17,14.5) {$y$};
        \node [label] (ypp) at (23,-4.5) {$y''$};

@@

        \node [label] (a) at (2,8.5) {$x'$};
        \node [label] (a) at (14,14.5) {$y'$};

@@

  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_2)-edge}]
    First, according to Lemma~\ref{lemma:at most one edge monotone},
    there exists at most one $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
    Suppose now, aiming at a contradiction, that there exists
    exactly one $(\nu_1, \nu_2)$-edge $e = (a, a')$ in $\mathcal{M}$.
    Under this assumption, we show that $\mathcal{M}$ does not contain any $(\nu_1, \nu_1)$-edge.
    Indeed, suppose, aiming at a contradiction,
    that there exists a $(\nu_1, \nu_1)$-edge $(x, x')$ in $\mathcal{M}$.
    Since $\mathcal{M}$ is containment-free,
    we must have $x < a$.
    Then it follows that $\mu(x) < \mu(a)$ (since $\nu_1$ is increasing) and
    $\mu(x') > \mu(a')$ (since $\nu_1$ is above $\nu_2$),
    which contradicts our hypothesis about $\mathcal{M}$.
    Hence $\mathcal{M}$ does not contain any $(\nu_1, \nu_1)$-edge.
    Furthermore, since $\nu_1$ is increasing and
    there exists only one $(\nu_1, \nu_2)$-edge $(a, a')$ in $\mathcal{M}$,
    we conclude that
    $\mathcal{M}$ contains $N_1-1$ pairwise crossing $(\nu_1, \nu'_1)$-edges.
    But $|\nu_1| = |\nu'_1| = N_1$ and $\mathcal{M}$ is perfect,
    and hence there exists a position, say $y$, in $\nu'_1$ that is
    not involved in a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
    We rule out this configuration by considering two cases:
    \begin{itemize}
      \item Position $y$ is involved in a $(\nu_2, \nu'_1)$-edge,
      say $(y', y)$. Then, we have $\mu(a) > \mu(y')$ (since $\nu_1$ is above $\nu_2$)
      and $\mu(a') < \mu(y)$ (since $\nu_2$ is below $\nu'_1$),
      which contradicts our hypothesis about $\mathcal{M}$.
      See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 1}.
      \item Position $y$ is not involved in a $(\nu_2, \nu'_1)$-edge.
      Since $\mathcal{M}$ contains $N_1-1$ pairwise crossing $(\nu_1, \nu'_1)$-edges
      in $\mathcal{M}$ and $|\nu'_1| = N_1$, there exists an edge $(y, y'')$ in $\mathcal{M}$
      that is neither a $(\nu_1, \nu'_1)$-edge nor a $(\nu_2, \nu'_1)$-edge
      nor a $(\nu'_1, \nu'_1)$-edge.
      In other words, $y'' > 2N_1 + N_2$.
      Now, for any $(\nu_1, \nu'_1)$-edge $(x, x')$ in $\mathcal{M}$, we
      have $\mu(x) < \mu(y)$ (since $\nu_1$ is below $\nu'_1$)
      and $\mu(x') > \mu(y'')$ (since $\nu'_1$ is above any other gadget),
      which contradicts our hypothesis about $\mathcal{M}$.
      See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 2}.
    \end{itemize}
    \qed
  \end{proof}

>>>>

  It is immediate that $\mu$ can be constructed in polynomial-time in $n$ and $k$.
  % We claim that $\sigma$ occurs in $\pi$ if and only if
  % there exists a containment-free perfect matching
  % $\mathcal{M}$ in $K_\mu$ such that
  % for any two disctinct edges
  % $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  % we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
  It can be shown (see Appendix) that
  that $\sigma$ occurs in $\pi$ if and only if
  there exists a containment-free perfect matching
  $\mathcal{M}$ in $K_\mu$ such that
  for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
  \qed
\end{proof}

@@

          \node [label] (y) at (17,14.5) {$b$};
          \node [label] (yp) at (10,-1.5) {$b'$};

@@

        \node [label] (y) at (17,14.5) {$b$};
        \node [label] (ypp) at (23,-4.5) {$b''$};

@@

        \node [label] (a) at (2,8.5) {$x$};
        \node [label] (a) at (14,14.5) {$x'$};

@@

  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_2)-edge}]
    First, according to Lemma~\ref{lemma:at most one edge monotone},
    there exists at most one $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
    Suppose now, aiming at a contradiction, that there exists
    exactly one $(\nu_1, \nu_2)$-edge $e = (a, a')$ in $\mathcal{M}$.
    Under this assumption, we show that $\mathcal{M}$ does not contain any $(\nu_1, \nu_1)$-edge.
    Indeed, suppose, aiming at a contradiction,
    that there exists a $(\nu_1, \nu_1)$-edge $(b, b')$ in $\mathcal{M}$.
    Since $\mathcal{M}$ is containment-free,
    we must have $b < a$.
    Then it follows that $\mu(b) < \mu(a)$ (since $\nu_1$ is increasing) and
    $\mu(b') > \mu(a')$ (since $\nu_1$ is above $\nu_2$),
    which contradicts our hypothesis about $\mathcal{M}$.
    Hence $\mathcal{M}$ does not contain any $(\nu_1, \nu_1)$-edge.
    Furthermore, since $\nu_1$ is increasing and
    there exists only one $(\nu_1, \nu_2)$-edge $(a, a')$ in $\mathcal{M}$,
    we conclude that
    $\mathcal{M}$ contains $N_1-1$ pairwise crossing $(\nu_1, \nu'_1)$-edges.
    But $|\nu_1| = |\nu'_1| = N_1$ and $\mathcal{M}$ is perfect,
    and hence there exists a position, say $b$, in $\nu'_1$ that is
    not involved in a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
    We rule out this configuration by considering two cases:
    \begin{itemize}
      \item Position $y$ is involved in a $(\nu_2, \nu'_1)$-edge,
      say $(b', b)$.
      Then, we have $\mu(a) > \mu(b')$ (since $\nu_1$ is above $\nu_2$)
      and $\mu(a') < \mu(b)$ (since $\nu_2$ is below $\nu'_1$),
      which contradicts our hypothesis about $\mathcal{M}$.
      See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 1}.
      \item Position $b$ is not involved in a $(\nu_2, \nu'_1)$-edge.
      Since $\mathcal{M}$ contains $N_1-1$ pairwise crossing $(\nu_1, \nu'_1)$-edges
      in $\mathcal{M}$ and $|\nu'_1| = N_1$, there exists an edge
      $(b, b'')$ in $\mathcal{M}$
      that is neither a $(\nu_1, \nu'_1)$-edge nor a $(\nu_2, \nu'_1)$-edge
      nor a $(\nu'_1, \nu'_1)$-edge.
      In other words, $y'' > 2N_1 + N_2$.
      Now, for any $(\nu_1, \nu'_1)$-edge $(x, x')$ in $\mathcal{M}$, we
      have $\mu(x) < \mu(b)$ (since $\nu_1$ is below $\nu'_1$)
      and $\mu(x') > \mu(b'')$ (since $\nu'_1$ is above any other gadget),
      which contradicts our hypothesis about $\mathcal{M}$.
      See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 2}.
    \end{itemize}
    \qed
  \end{proof}
